# Making our package

## Multiple files.

A good practice is to make many .R files rather than putting all your code into one huge one. These files should also be named in a way where the contents are going to be clear.

***Exercise:*** Make a new file in `R/` which has your code to make a new instance of `mysco`. Document the function so people know how to use it.

What you should be aiming for is being able to do this:

```{r,eval=F}
library(SCAP)

pbmc.data <- readRDS("PBMC_data.rds")
pbmc <- CreateMySCO(pbmc.data)
```

Put your `show` method in to the package.

Now add your `CalcMitoPct`,`MakeQCPlots(pbmc)`, and `FilterData` into your package.


So you should be able to run these lines:

```{r,eval=F}
library(SCAP)

pbmc.data <- readRDS("PBMC_data.rds")
pbmc <- CreateMySCO(pbmc.data)
pbmc <- CalcMitoPct(pbmc,"^MT-")
MakeQCPlots(pbmc)
pbmc <- FilterData(pbmc,sub=features > 200 & features < 2500 & perc.mt < 5,min.cells=3)
pbmc
```

Easy right? :)

When you're happy that your package is working so far, push it up to Github.

